---
title: "**Seeing the Other Side: How Local Party Images Moderate Affective Polarization**"
author: |
  Chaoyue Wang^[Chaoyue R. Wang (<chyrwang@gmail.com>, `86-177-2387-5369`) is a senior student of Philosophy, Politics and Economics at Yuanpei College of Peking University, Beijing, China 100871. Data, codes, and other documents necessary to reproduce the results reported in this study are openly available at <https://github.com/Raphaellie/Local-Party-Images>.]
abstract: |
  Social sorting, the situation where social groups become strongly interconnected with political parties, has been examined as a contributing factor to affective polarization in American politics. By cultivating party images represented through prominent social identities, this process instills group tensions into party politics. Yet few existing studies takes advantage of local variations of partisan composition to investigate whether the extent to which current political polarization are affected by “parties in our head”. This research builds an index for party images for each congressional district using 2016 and 2020 Cooperative Election Studies, and links these local measures with 2020 American Election Studies where detailed indicators of polarization behavior are included. I find that the further the racial images of local partisans deviate from national stereotypes, the lower level of affective polarization is observed among the district’s respondents. The deviation of local party images also undermines the influence group feelings have in people’s polarization behavior. Heterogeneities by party and racial group are discussed.
format: 
  pdf:
    documentclass: article
    pdf-engine: lualatex
    fig_caption: yes
    fontsize: 12pt
    toc: false
    indent: true
    number-sections: true
    fontfamily: crimson
    linestretch: 1.5
    geometry: margin = 1.2in
    include-before: 
      - \captionsetup{labelfont = bf, font = small} 
      - \AtBeginEnvironment{tabular}{\small}
      - \AtBeginEnvironment{tablenotes}{\small}
      - \urlstyle{tt}
    header-includes: 
      - \usepackage[scale = 0.75]{sourcecodepro}
      - \usepackage{etoolbox}
    fig-cap-location: bottom
    tbl-cap-location: top
    fig-pos: tb
    tbl-pos: tb
    link-citations: true
    keep-tex: true
    colorlinks: true
    citecolor: NavyBlue
    linkcolor: NavyBlue
    urlcolor: SteelBlue4
    template-partials:
      - before-body.tex
  html:
    theme: [journal,custom.scss] # lumen journal flatly cosmo
    fig-dpi: 250
    number-sections: true
    code-tools: true
    toc: true
    toc-title: Overview
    toc-location: left
    title-block-banner: '#f0f1f2'
    title-block-banner-color: '#10378c'
    comments:
      hypothesis:
        theme: classic
# csl: ajps.csl
editor: visual
execute:
  echo: false
  warning: false
  cache: false
date: last-modified
date-format: long
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

```{r}
#| label: setup
library(ggplot2)
library(sjPlot)
library(modelsummary)
library(patchwork)

load('final.Rdata')
options(modelsummary_format_numeric_latex = "latex")
options("modelsummary_stars_note" = FALSE)
gm <- tibble::tribble(
  ~raw,        ~clean,          ~fmt,
  "nobs",      "Observations",             0,
  "r.squared", "R squared", 3)
table.output <- 'kableExtra'
```

# Introduction

Their ideological differences aside, the Republican and Democratic Parties in the United States represent two racially divergent faces of the country [@mason2018; @egan2020]. The passage of prominent civil rights legislation in the 1960s activated a profound realignment, sorting blacks to the Democratic Party and racially conservative whites to the GOP [@carmines1990; @valentino2005]. This shakeup of the race-party landscape, combined with the overall diversification of the U.S. population in the past half a century [@hajnal2014], resulted in a sharp contrast between the racial makeups of the two major parties: while the Republican Party remains principally white, the Democratic Party embodies a racially diverse America [@mason2016; @egan2020; @zhirkov2022].

Such contrast has led scholars to construe the origins and consequences of mass partisanship through a perspective that centers on the role of social identities. Breaking from the instrumental tradition that largely reduces partisanship to a "running tally" of police preferences [@fiorina1981; @downs1957], the group-oriented approach to American partisanship views such identification as a expressive product of one's social thinking [@green2002; @mason2018; @kane2021; @zhirkov2022]. It contends that calling the distinct images of Republicans and Democrats to mind, individuals will decide their partisanship based upon which party most closely resembles their own social images [@green2002; @mason2018], or the images of those groups they favor [@kane2021]. In this sense, the political self is primarily an extension of one's other social selves. Scholars taking the groups approach have found that during the last decades, there is an incremental overlap between group affect and partisan affect [@zhirkov2022]. Also, a closer match of one's group identities with their party's image is associated with strengthened partisan identity and the escalation of affective polarization between co-partisans and out-partisans [@mason2018a; @mason2016].

As is evident in the groups approach, party images[^1], the mental schema that communicates what the Republican and Democratic Parties look like in terms of social groups [@green2002; @ahler2018; @zhirkov2022], play a pivotal role in connecting group identities to political identification. Though such images are so pervasive and entrenched that party and race are almost "inseparable" in the public mind [@westwood2020], variation of them still exists. To check the extent to which "the parties in our heads" moderate the linkage between group feelings and partisan affect, existing studies have measured and manipulated subjects' perceptions of party images in experimental settings. Researchers find a decline of partisan affective polarization when individuals possess weaker knowledge about what party goes with what groups [@zhirkov2022], or when their prior perceptions of a party's social makeup are challenged [@ahler2018]. Yet to the best of my knowledge, there has not been any attempt made how the variation of party images that derive from contextual, local conditions may enter into the connection between group affect and political polarization.

[^1]: This same concept has been called in various terms across studies. For example, whereas @kane2021 adopt the phrase "group-party alignment knowledge", @ahler2018 and @zhirkov2022 respectively settle on "partisan prototypes" and "race-party schemas". This paper uses "party images", the term originally developed by @green2002 [p.8] in reference to group compositions of Democrats, Republicans, and Independents.

With the consistent and structural influence of local context in mind [@campbell2006; @hopkins2009; @newman2015; @wong2010], this study advances the group-oriented research of American partisanship by looking into how the variation of party images among congressional districts moderates the connection between group thinking and affective polarization. Pooling the 2016 and 2020 Cooperative Election Study (CES) data, I measure the racialized party images of the Democratic and Republican partisans for every congressional district. I then match this district-level data with respondents in 2020 American National Election Study (ANES) in which affective polarization is recorded for each individual.

I find that party images, expressed in terms of white imagery, does not produce standalone effects on affective polarization. But when put in comparison, the contrast between the racial images of the two parties significantly influences the affective polarization level of a district. Specifically, for the partisans whose group memberships rightly aligns with their party, affective polarization climbs up as the racial images of the two parties diverge further; for those embodying a ambivalent combination of racial and partisan identities, the increase in party image contrast results in a decline in affective polarization. Finally, I show that how much citizens link their racial affect to partisan feelings is moderated by the degree to which Democrats and Republicans in their district look (un)like each other, thereby validating the theoretical role of divergent party images as an essential hinge in current racial-political divides.

# Party Images and Affective Polarization

# Empirical Strategy

## Measure District-Level Party Images

```{r}
#| label: fig-map
#| fig-cap: "**Geograhpic Variation of Party Imageries at the Level of Congressional Districts.** States are resized to be in proportion to their population sizes, and districts are located within the states to approximately match their actual locations. This cartogram is created by the Daily Kos team (<https://www.dailykos.com>)."
#| fig-width: 9
#| fig-height: 8

fig.map
```

```{r}
#| label: fig-hist
#| fig-cap: "**Distribution of Party Imageries and Their Gap among Congressional Districts.** The first two panels shows the distribution of the percentages of non-Hispanic whites among the Republican and Democratic partisans in a congressional district. The last panel, capturing the contrast of the white imagery between Republicans and Democrats, shows the distribution of the gap between the two percentages."
#| fig-width: 10
#| fig-asp: 0.33

fig.hist
```

```{r}
#| label: fig-baseline
#| fig-cap: "**The Estimated Effect of Party Imageries on Affective Polarization by Different Types of Partisans.** Panels indicates three subsamples of partisans in 2020 ANES. The labels on the horizontal axis shows the party imagery of what political group is of interest in a regression. A single point represents the estimated effect of the district-level white imagery of a political group on the affective polarization of the partisans in a congressional district, and the range shows its 95% confidence interval. Significant esimtates (p < 0.05) are colored in green and marked in bold text. White percentage of a district's population is controlled for. "
#| fig-width: 10
#| fig-asp: 0.33

fig.baseline
```

```{r}
#| label: fig-decompose
#| fig-cap: "**The Estimated Effect of Party Imagery on Affective Polarization, Broken Down by In-Group and Out-Group Feelings.** Panels indicates three outcome variables pertaining to affective polarization, and the labels on the vertical axis shows the partisan subsample of 2020 ANES upon whom a regression is based. An individual point represents the estimated effect that district-level contrast of party imagery has on the outcome variable, and the range shows its 95% confidence interval. Significant esimtates (p < 0.05) are colored in red and marked in bold text. White percentage of a district's population is controlled for."
#| fig-width: 9
#| fig-asp: 0.4

fig.decompose
```

```{r}
#| label: tbl-mechanism
#| tbl-cap: "Contrast of Party Images Moderates the Effect of White Feelings on Affective Polarization"

modelsummary(
  models, 
  stars = TRUE, output = table.output,
  threeparttable = TRUE,
  gof_map = gm,
  coef_map = c(
    "ft_white" = "White FT",
    "gap" = 'Party Imagery Contrast',
    "ft_white:gap" = 'White FT × Party Imagery Contrast'),
  notes = "Note: All respondents in 2020 ANES. FT means the respondent's feeling thermometer score of the group on a 100-point scale, higher values indicating more warmer feelings. The term Party Imagery Contrast refers to the contrast between Republicans and Democrats in terms of their white imagery. Robust standard errors in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001" )
```

```{r}
#| label: fig-mechanism
#| fig-cap: "**Greater Contrast of Party Images Accentuates the Influence of White Feelings on Voter's Affective Perception of Political Parties.** Linear predicted values based upon the previous OLS analyses with interaction terms between white FT and party FT. Panels reponds to the three outcome variables. Within each panel, the two ribbons plot how affective polarization changes as feelings toward whites become warmer at two levels of party imagery contrast. The bands show the 95% confidcence internals. "
#| fig-width: 9
#| fig-height: 4

fig.mechanism
```

\small

# References
