---
title: "See the Other Side"
subtitle: How Local Party Images Moderate Affective Polarization
author: "Chaoyue R. Wang"
institute: Peking University
fig-dpi: 200
format: 
  revealjs:
    theme: simple
    highlight-style: github
    slide-number: c/t
execute: 
  echo: false
  cache: false
editor: visual
---

# Theory

## Race and Parties in U.S. Politics

-   In studying US partisanship, scholars often invoke a groups approach, wherein individuals' group memberships affect their identification with the Democratic and Republican Parties.

-   Reasonably accurate knowledge about how social groups align with political parties is an essential component that activates the link between group affect and party attachments.

# Results

```{r}
#| label: setup
#| wanring: false
library(ggplot2)
library(sjPlot)
library(modelsummary)
library(patchwork)

load('final.Rdata')
options(modelsummary_format_numeric_latex = "latex")
options("modelsummary_stars_note" = FALSE)
gm <- tibble::tribble(
  ~raw,        ~clean,          ~fmt,
  "nobs",      "Observations",             0,
  "r.squared", "R squared", 3)
table.output <- 'kableExtra'
```

## Affective Polarization and Party Images

```{r}
#| label: fig-baseline
#| fig-cap: "**The Estimated Effect of Party Imageries on Affective Polarization by Different Types of Partisans.**  "
#| fig-width: 8
#| fig-asp: 0.43

fig.baseline
```

## The Role of Party Image Contrast

```{r}
#| label: fig-decompose
#| fig-cap: "**The Estimated Effect of Party Imagery on Affective Polarization, Broken Down by In-Group and Out-Group Feelings.** "
#| fig-width: 8
#| fig-asp: 0.4

fig.decompose
```

## Influence of Racial Feelings Weakened

```{r}
#| label: fig-mechanism
#| fig-cap: "**Greater Contrast of Party Images Accentuates the Influence of White Feelings on Voter's Affective Perception of Political Parties.**"
#| fig-width: 8
#| fig-asp: 0.45

fig.mechanism
```
